---
title: "Linear Regression Models"
author: "Herong Wang"
date: "9/28/2022"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
#Sys.setenv(R_USER ="C:/Users/15340/Documents")
setwd("C:/Users/herongw/Desktop/Kelly_RA/Projects/depressive_DNAm_HRS/Brittant_original/code_dataset")


```

```{r}
#install.packages("nnet") # for multinomial regression
#install.packages("tidyverse")
library(tidyverse)
library(car)
library(mediation)
library(meta)
library(metasens)
library(nnet)
library(devtools)
#devtools::install_github("BS1125/CMAverse")
library(CMAverse)
library(ggplot2)
if(!require(devtools)) install.packages("devtools")
#devtools::install_github("kassambara/ggpubr")
library(ggpubr)
library(lmtest)

load("analytic_sample_3793.RData")


```

## check collinearity between composition of cells

```{r}
cell_perc <- complete %>% 
  select(gran, PMONO, PLYMP)

cor_matrix <- rcorr(as.matrix(cell_perc))
cor_matrix
cor.test(cell_perc$gran, cell_perc$PLYMP) ## percent of granulocytes is highly correlated with                                                              lymphocytes

cor.test(complete$R13SHLT, complete$R13CONDE)

load("included_3915.RData")
cor.test(wave13_included$gran, wave13_included$PLYMP)
```

## Unadjusted model

```{r}
model_unadj <- lm(GRIM_residuals ~ depression, data = complete)

plot(model_unadj)
summary(model_unadj)
confint(model_unadj)
```

## Primary model- age, gender, race/ethnicity, highest degree, marital status

```{r}

model_prim <- lm(GRIM_residuals ~ depression + age + GENDER + race_ethnicity + DEGREE_collapsed + PMARST_collapsed + gran + PMONO  , data = complete)

plot(model_prim)
summary(model_prim)
confint(model_prim)

```

## Secondary model- including health behaviors

```{r}

model_hlth <- lm(GRIM_residuals ~ depression + age + GENDER + race_ethnicity + DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO , data = complete)

plot(model_hlth)
summary(model_hlth)
confint(model_hlth)

### check interaction ###

model_hlth_inter_race <- lm(GRIM_residuals ~ depression + age + GENDER + race_ethnicity + DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO + depression*race_ethnicity , data = complete)
summary(model_hlth_inter_race)

model_hlth_inter_sex <- lm(GRIM_residuals ~ depression + age + GENDER + race_ethnicity + DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO + depression*GENDER , data = complete)
summary(model_hlth_inter_sex)

model_hlth_inter_sex_race <- lm(GRIM_residuals ~ depression + age + GENDER + race_ethnicity + DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO + depression*GENDER*race_ethnicity , data = complete)
summary(model_hlth_inter_sex_race)

## check multicollinearity ##
car::vif(model_hlth)

# Checking which variable is responsible for large change

model_hlth1 <- lm(GRIM_residuals ~ depression + age + GENDER + RACE + DEGREE_collapsed + PMARST_collapsed + R13CONDE, data = complete)
summary(model_hlth1) # coefficient is 0.51

table(complete$R13CONDE)

model_hlth2 <- lm(GRIM_residuals ~ depression + age + GENDER + RACE + DEGREE_collapsed + PMARST_collapsed + R13DRINKD, data = complete)
summary(model_hlth2) # coefficient is 1.21

model_hlth3 <- lm(GRIM_residuals ~ depression + age + GENDER + RACE + DEGREE_collapsed + PMARST_collapsed + smoke, data = complete)
summary(model_hlth3) # coefficient is 1.21

model_hlth4 <- lm(GRIM_residuals ~ depression + age + GENDER + RACE + DEGREE_collapsed + PMARST_collapsed + gran, data = COMPLETE)
summary(model_hlth4) # coefficient is 0.92

model_hlth5 <- lm(GRIM_residuals ~ depression + age + GENDER + RACE + DEGREE_collapsed + PMARST_collapsed + PMONO, data = complete) # coefficient is 1.20
summary(model_hlth5)
```

## likelihood test of the model fit between unadjusted & primary & secondary models
### Primary analysis
```{r}
## unadjusted and primary models
lrtest(model_prim, model_unadj)

## primary model and health behavior model (secondary model)
lrtest(model_hlth, model_prim)
```

## Sensitivity analyses

### Models with CES-D score continuous

```{r}

model_cesd_con <- lm(GRIM_residuals ~ R13CESD, data = complete)

plot(model_cesd_con)
summary(model_cesd_con)
confint(model_cesd_con)

# Primary adjusted model

model_cesd_con2 <- lm(GRIM_residuals ~ R13CESD + age + GENDER + race_ethnicity + DEGREE_collapsed + PMARST_collapsed + gran + PMONO , data = complete)

plot(model_cesd_con2)
summary(model_cesd_con2)
confint(model_cesd_con2)

# Secondary adjusted model 

model_cesd_con3 <- lm(GRIM_residuals ~ R13CESD + age + GENDER + race_ethnicity + DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO , data = complete)

plot(model_cesd_con3)
summary(model_cesd_con3)
confint(model_cesd_con3)

```

# Model with any CES-D score greater than 0 as exposed

```{r}

model_cesd_any <- lm(GRIM_residuals ~ any_dep, data = complete)

plot(model_cesd_any)
summary(model_cesd_any)
confint(model_cesd_any)

# Primary adjusted model

model_cesd_any2 <- lm(GRIM_residuals ~ any_dep + age + GENDER + race_ethnicity + DEGREE_collapsed + PMARST_collapsed + gran + PMONO , data = complete)

plot(model_cesd_any2)
summary(model_cesd_any2)
confint(model_cesd_any2)

# Secondary adjusted model 

model_cesd_any3 <- lm(GRIM_residuals ~ any_dep + age + GENDER + race_ethnicity + DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO , data = complete)

plot(model_cesd_any3)
summary(model_cesd_any3)
confint(model_cesd_any3)

```

# Model with Horvath clock residuals as the outcome variable

```{r}
model_horvath <- lm(HORVATH_residuals ~ depression, data = complete)

plot(model_horvath)
summary(model_horvath)
confint(model_horvath)

# Primary adjusted model

model_horvath2 <- lm(HORVATH_residuals ~ depression + age + GENDER + race_ethnicity + DEGREE_collapsed + PMARST_collapsed + gran + PMONO , data = complete)

plot(model_horvath2)
summary(model_horvath2)
confint(model_horvath2)

# Secondary adjusted model 

model_horvath3 <- lm(HORVATH_residuals ~ depression + age + GENDER + race_ethnicity + DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO, data = complete)

plot(model_horvath3)
summary(model_horvath3)
confint(model_horvath3)

```

# Model with LEVINE clock residuals as the outcome variable

```{r}
model_LEVINE <- lm(LEVINE_residuals ~ depression, data = complete)

plot(model_LEVINE)
summary(model_LEVINE)
confint(model_LEVINE)

# Primary adjusted model

model_LEVINE2 <- lm(LEVINE_residuals ~ depression + age + GENDER + race_ethnicity + DEGREE_collapsed + PMARST_collapsed + gran + PMONO , data = complete)

plot(model_LEVINE2)
summary(model_LEVINE2)
confint(model_LEVINE2)

# Secondary adjusted model 

model_LEVINE3 <- lm(LEVINE_residuals ~ depression + age + GENDER + race_ethnicity + DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO, data = complete)

plot(model_LEVINE3)
summary(model_LEVINE3)
confint(model_LEVINE3)

```

# Model with MPOA clock residuals as the outcome variable

```{r}
model_MPOA <- lm(MPOA_residuals ~ depression, data = complete)

plot(model_MPOA)
summary(model_MPOA)
confint(model_MPOA)

# Primary adjusted model

model_MPOA2 <- lm(MPOA_residuals ~ depression + age + GENDER + race_ethnicity + DEGREE_collapsed + PMARST_collapsed + gran + PMONO , data = complete)

plot(model_MPOA2)
summary(model_MPOA2)
confint(model_MPOA2)

# Secondary adjusted model 

model_MPOA3 <- lm(MPOA_residuals ~ depression + age + GENDER + race_ethnicity + DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO, data = complete)

plot(model_MPOA3)
summary(model_MPOA3)
confint(model_MPOA3)

```

# Model with HANNUM clock residuals as the outcome variable

```{r}
model_HANNUM <- lm(HANNUM_residuals ~ depression, data = complete)

plot(model_HANNUM)
summary(model_HANNUM)
confint(model_HANNUM)

# Primary adjusted model

model_HANNUM2 <- lm(HANNUM_residuals ~ depression + age + GENDER + race_ethnicity + DEGREE_collapsed + PMARST_collapsed + gran + PMONO, data = complete)

plot(model_HANNUM2)
summary(model_HANNUM2)
confint(model_HANNUM2)

# Secondary adjusted model 

model_HANNUM3 <- lm(HANNUM_residuals ~ depression + age + GENDER + race_ethnicity + DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO, data = complete)

plot(model_HANNUM3)
summary(model_HANNUM3)
confint(model_HANNUM3)

```

# Model adding physical activity and childhood poverty

```{r}
model_act <- lm(GRIM_residuals ~ depression + age + GENDER + race_ethnicity + DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO + R13ACT, data = complete)
summary(model_act)
confint(model_act)

model_pov <- lm(GRIM_residuals ~ depression + age + GENDER + race_ethnicity + DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO + as.factor(poverty), data = complete)
summary(model_pov)
confint(model_pov)

model_sens <- lm(GRIM_residuals ~ depression + age + GENDER + race_ethnicity + DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO + R13ACT + as.factor(poverty), data = complete)

plot(model_sens)
summary(model_sens)
confint(model_sens)

## CALCULATE AIC ##
AIC(model_hlth)
AIC(model_act)
AIC(model_pov)
AIC(model_sens)

## ANOVA test for model fit ##
anova(model_hlth, model_pov, test = "F")
anova(model_pov, model_sens, test = "F")
anova(model_hlth, model_sens, test = "F")

## likelihood test for model fit ##
lrtest(model_pov, model_hlth)
lrtest(model_sens, model_pov)
lrtest(model_sens, model_hlth)

```

## race-specific regression

```{r dev='jpeg'}

complete_white = complete %>% 
  filter(race_ethnicity == "Non-Hispanic White")

model_white_unad = lm(GRIM_residuals ~ depression, data = complete_white)
summary(model_white_unad)
confint(model_white_unad)

model_white_prim <- lm(GRIM_residuals ~ depression + age + GENDER  + DEGREE_collapsed + PMARST_collapsed + gran + PMONO  , data = complete_white)
summary(model_white_prim)
confint(model_white_prim)

model_white_hlth <- lm(GRIM_residuals ~ depression + age + GENDER +  DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO , data = complete_white)
summary(model_white_hlth)
confint(model_white_hlth)


complete_black = complete %>% 
  filter(race_ethnicity == "Non-Hispanic Black")

model_black_unad = lm(GRIM_residuals ~ depression, data = complete_black)
summary(model_black_unad)
confint(model_black_unad)

model_black_prim <- lm(GRIM_residuals ~ depression + age + GENDER  + DEGREE_collapsed + PMARST_collapsed + gran + PMONO  , data = complete_black)
summary(model_black_prim)
confint(model_black_prim)

model_black_hlth <- lm(GRIM_residuals ~ depression + age + GENDER +  DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO  , data = complete_black)
summary(model_black_hlth)
confint(model_black_hlth)

complete_other = complete %>% 
  filter(race_ethnicity == "Non-Hispanic Other")

model_other_unad = lm(GRIM_residuals ~ depression, data = complete_other)
summary(model_other_unad)
confint(model_other_unad)

model_other_prim <- lm(GRIM_residuals ~ depression + age + GENDER  + DEGREE_collapsed + PMARST_collapsed + gran + PMONO  , data = complete_other)
summary(model_other_prim)
confint(model_other_prim)

model_other_hlth <- lm(GRIM_residuals ~ depression + age + GENDER +  DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO  , data = complete_other)
summary(model_other_hlth)
confint(model_other_hlth)

complete_hispanic = complete %>% 
  filter(race_ethnicity == "Hispanic")

model_his_unad = lm(GRIM_residuals ~ depression, data = complete_hispanic)
summary(model_his_unad)
confint(model_his_unad)

model_his_prim <- lm(GRIM_residuals ~ depression + age + GENDER  + DEGREE_collapsed + PMARST_collapsed + gran + PMONO  , data = complete_hispanic)
summary(model_his_prim)
confint(model_his_prim)

model_his_hlth <- lm(GRIM_residuals ~ depression + age + GENDER +  DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO  , data = complete_hispanic)
summary(model_his_hlth)
confint(model_his_hlth)

library(esc)
library(tidyverse)

# meta analysis ##
summary(model_white_hlth)
white_ex_mean <- model_white_hlth$coefficients[1]+model_white_hlth$coefficients[2]
white_ex_sd <- summary(model_white_hlth)$coefficients[2,2]
white_ex_n <- 293
white_unex_mean <- model_white_hlth$coefficients[1]
white_unex_sd <- summary(model_white_hlth)$coefficients[1,2]
white_unex_n <- 2246
group <- "Non-Hispanic White"
white <- data.frame(group, white_ex_mean,white_ex_sd,white_ex_n,white_unex_mean,white_unex_sd,white_unex_n)

colnames(white) = c('group', "ex_mean","ex_sd","ex_n","unex_mean","unex_sd","unex_n")

summary(model_black_hlth)
black_ex_mean <- model_black_hlth$coefficients[1]+model_black_hlth$coefficients[2]
black_ex_sd <- summary(model_black_hlth)$coefficients[2,2]
black_ex_n <- 115
black_unex_mean <- model_black_hlth$coefficients[1]
black_unex_sd <- summary(model_black_hlth)$coefficients[1,2]
black_unex_n <- 518

table(complete_black$depression)
group <- "Non-Hispanic Black"
black <- data.frame(group, black_ex_mean,black_ex_sd,black_ex_n,black_unex_mean,black_unex_sd,black_unex_n)

colnames(black) = c('group', "ex_mean","ex_sd","ex_n","unex_mean","unex_sd","unex_n")

other_ex_mean <- model_other_hlth$coefficients[1]+model_other_hlth$coefficients[2]
other_ex_sd <- summary(model_other_hlth)$coefficients[2,2]
other_ex_n <- 22
other_unex_mean <- model_other_hlth$coefficients[1]
other_unex_sd <- summary(model_other_hlth)$coefficients[1,2]
other_unex_n <- 92
group <- "Non-Hispanic Other"
other <- data.frame(group, other_ex_mean,other_ex_sd,other_ex_n,other_unex_mean,other_unex_sd,other_unex_n)
colnames(other) = c('group', "ex_mean","ex_sd","ex_n","unex_mean","unex_sd","unex_n")

his_ex_mean <- model_his_hlth$coefficients[1]+model_his_hlth$coefficients[2]
his_ex_sd <- summary(model_his_hlth)$coefficients[2,2]
his_ex_n <- 114
his_unex_mean <- model_his_hlth$coefficients[1]
his_unex_sd <- summary(model_his_hlth)$coefficients[1,2]
his_unex_n <- 393
group <- "Hispanic"
hispanic <- data.frame(group, his_ex_mean,his_ex_sd,his_ex_n,his_unex_mean,his_unex_sd,his_unex_n)
colnames(hispanic) = c('group', "ex_mean","ex_sd","ex_n","unex_mean","unex_sd","unex_n")

table(complete_hispanic$depression)

meta <- rbind(white, black, other, hispanic)
rownames(meta) <- c("1", "2", "3", "4")


library(meta)
library(metasens)

meta_race =  metacont(ex_n, ex_mean, ex_sd, 
              unex_n, unex_mean, unex_sd,
              comb.fixed = T, comb.random = T, studlab = group,
              data = meta, sm = "SMD") 
meta_race

forest_race = forest(meta_race, print.I2 = F, print.I2.ci = F, print.tau2 = F, print.pval.Q = F,
       leftcols = c("studlab",  "effect.ci"), leftlabs = c("Strata", "Effect Estimate (95% CI)"),
       rightcols = c("w.fixed", "w.random"), xlim = c(-1,1))


#### do not use metacont() here
######## metacont() is used in raw data instead of pre-calculated effect size
##################### for pre-calculated effect size, we can use metagen()

group = "White"
effect.size = model_white_hlth$coefficients[2]
effect.size.se <- summary(model_white_hlth)$coefficients[2,2]
white <- data.frame(group, effect.size, effect.size.se)
colnames(white) = c('group', "effect.size", "effect.size.sd")

group = "Black"
effect.size = model_black_hlth$coefficients[2]
effect.size.se <- summary(model_black_hlth)$coefficients[2,2]
black <- data.frame(group, effect.size, effect.size.se)
colnames(black) = c('group', "effect.size", "effect.size.sd")

group = "Other"
effect.size = model_other_hlth$coefficients[2]
effect.size.se <- summary(model_other_hlth)$coefficients[2,2]
other <- data.frame(group, effect.size, effect.size.se)
colnames(other) = c('group', "effect.size", "effect.size.sd")

group = "Hispanic"
effect.size = model_his_hlth$coefficients[2]
effect.size.se <- summary(model_his_hlth)$coefficients[2,2]
his <- data.frame(group, effect.size, effect.size.se)
colnames(his) = c('group', "effect.size", "effect.size.sd")

meta_race <- rbind(white, black, his)
rownames(meta_race) <- c("1", "2", "3")

meta.race <- metagen(TE = effect.size,
                 seTE = effect.size.sd,
                 studlab = group,
                 data = meta_race,
                 sm = "MD",
                 fixed = T,
                 random = F,
                 method.tau = "REML",
                 hakn = TRUE)

summary(meta.race)

forest_race = forest(meta.race,print.I2 = F, print.I2.ci = F, print.tau2 = F, print.pval.Q = T, 
       leftcols = c("studlab", "effect.ci"), leftlabs = c("Strata", "Effect Estimate (95% CI)"),
       rightcols = c("w.fixed"), xlim = c(-2,2))
```


## sex specific regression

```{r}
complete_male <- complete %>% 
  filter(GENDER == "Male")

model_male_unad = lm(GRIM_residuals ~ depression, data = complete_male)
summary(model_male_unad)
confint(model_male_unad)

model_male_prim <- lm(GRIM_residuals ~ depression + age + race_ethnicity  + DEGREE_collapsed + PMARST_collapsed + gran + PMONO, data = complete_male)
summary(model_male_prim)
confint(model_male_prim)

model_male_hlth <- lm(GRIM_residuals ~ depression + age + race_ethnicity +  DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO  , data = complete_male)
summary(model_male_hlth)
confint(model_male_hlth)

complete_female <- complete %>% 
  filter(GENDER == "Female")

model_female_unad = lm(GRIM_residuals ~ depression, data = complete_female)
summary(model_female_unad)
confint(model_female_unad)

model_female_prim <- lm(GRIM_residuals ~ depression + age + race_ethnicity  + DEGREE_collapsed + PMARST_collapsed + gran + PMONO, data = complete_female)
summary(model_female_prim)
confint(model_female_prim)

model_female_hlth <- lm(GRIM_residuals ~ depression + age + race_ethnicity +  DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO  , data = complete_female)
summary(model_female_hlth)
confint(model_female_hlth)


## meta analysis ##
male_ex_mean <- mean(subset(complete,GENDER == "Male" & depression == "Elevated depressive symptoms")$GRIM_residuals)
male_ex_sd <- sd(subset(complete,GENDER == "Male" & depression == "Elevated depressive symptoms")$GRIM_residuals)
male_ex_n <- 170
male_unex_mean <- mean(subset(complete,GENDER == "Male" & depression == "Low or no depressive symptoms")$GRIM_residuals)
male_unex_sd <- sd(subset(complete,GENDER == "Male" & depression == "Low or no depressive symptoms")$GRIM_residuals)
male_unex_n <- 1408
group <- "Male"
male <- data.frame(group, male_ex_mean,male_ex_sd,male_ex_n,male_unex_mean,male_unex_sd,male_unex_n)
colnames(male) = c('group', "ex_mean","ex_sd","ex_n","unex_mean","unex_sd","unex_n")

female_ex_mean <- mean(subset(complete,GENDER == "Female" & depression == "Elevated depressive symptoms")$GRIM_residuals)
female_ex_sd <- sd(subset(complete,GENDER == "Female" & depression == "Elevated depressive symptoms")$GRIM_residuals)
female_ex_n <- 374
female_unex_mean <- mean(subset(complete,GENDER == "Female" & depression == "Low or no depressive symptoms")$GRIM_residuals)
female_unex_sd <- sd(subset(complete,GENDER == "Female" & depression == "Low or no depressive symptoms")$GRIM_residuals)
female_unex_n <- 1841
group <- "Female"
female <- data.frame(group, female_ex_mean, female_ex_sd, female_ex_n, female_unex_mean, female_unex_sd, female_unex_n)
colnames(female) = c('group', "ex_mean","ex_sd","ex_n","unex_mean","unex_sd","unex_n")

meta_sex <- rbind(male, female)
rownames(meta_sex) <- c("1", "2")

meta_r_sex =  metacont(ex_n, ex_mean, ex_sd, 
              unex_n, unex_mean, unex_sd,
              comb.fixed = T, comb.random =F, studlab = group,
              data = meta_sex, sm = "MD") 
summary(meta_r_sex)

forest_sex = forest(meta_r_sex,print.I2 = F, print.I2.ci = F, print.tau2 = F, print.pval.Q = F, 
       leftcols = c("studlab", "effect.ci"), leftlabs = c("Strata", "Effect Estimate (95% CI)"),
       rightcols = c("w.fixed", "w.random"), xlim = c(-1,1))

#### do not use metacont() here
######## metacont() is used in raw data instead of pre-calculated effect size
##################### for pre-calculated effect size, we can use metagen()
group = "Male"
effect.size = model_male_hlth$coefficients[2]
effect.size.se <- summary(model_male_hlth)$coefficients[2,2]
male <- data.frame(group, effect.size, effect.size.se)
colnames(male) = c('group', "effect.size", "effect.size.sd")

group = "Female"
effect.size = model_female_hlth$coefficients[2]
effect.size.se <- summary(model_female_hlth)$coefficients[2,2]
female <- data.frame(group, effect.size, effect.size.se)
colnames(female) = c('group', "effect.size", "effect.size.sd")

meta_sex <- rbind(male, female)
rownames(meta_sex) <- c("1", "2")

meta.sex <- metagen(TE = effect.size,
                 seTE = effect.size.sd,
                 studlab = group,
                 data = meta_sex,
                 sm = "MD",
                 fixed = T,
                 random = F,
                 method.tau = "REML",
                 hakn = TRUE)

summary(meta.sex)

forest_sex = forest(meta.sex,print.I2 = F, print.I2.ci = F, print.tau2 = F, print.pval.Q = T, 
       leftcols = c("studlab", "effect.ci"), leftlabs = c("Strata", "Effect Estimate (95% CI)"),
       rightcols = c("w.fixed"), xlim = c(-2,2))
```

## sex and race models and forest plot
```{r}
table(complete$GENDER,complete$race_ethnicity)

d = subset(complete, GENDER == "Male" & race_ethnicity == "Non-Hispanic White" )

white_male = lm(GRIM_residuals ~ depression + age  +  DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO  , data = subset(complete, GENDER == "Male" & race_ethnicity == "Non-Hispanic White" ))
summary(white_male)
confint(white_male)

white_female = lm(GRIM_residuals ~ depression + age  +  DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO  , data = subset(complete, GENDER == "Female" & race_ethnicity == "Non-Hispanic White" ))
summary(white_female)
confint(white_female)

black_male = lm(GRIM_residuals ~ depression + age  +  DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO  , data = subset(complete, GENDER == "Male" & race_ethnicity == "Non-Hispanic Black" ))
summary(black_male)
confint(black_male)

black_female = lm(GRIM_residuals ~ depression + age  +  DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO  , data = subset(complete, GENDER == "Female" & race_ethnicity == "Non-Hispanic Black" ))
summary(black_female)
confint(black_female)

other_male = lm(GRIM_residuals ~ depression + age  +  DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO  , data = subset(complete, GENDER == "Male" & race_ethnicity == "Non-Hispanic Other" ))
summary(other_male)
confint(other_male)

other_female = lm(GRIM_residuals ~ depression + age  +  DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO  , data = subset(complete, GENDER == "Female" & race_ethnicity == "Non-Hispanic Other" ))
summary(other_female)
confint(other_female)

hispanic_male = lm(GRIM_residuals ~ depression + age  +  DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO  , data = subset(complete, GENDER == "Male" & race_ethnicity == "Hispanic" ))
summary(hispanic_male)
confint(hispanic_male)

hispanic_female = lm(GRIM_residuals ~ depression + age  +  DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO  , data = subset(complete, GENDER == "Female" & race_ethnicity == "Hispanic" ))
summary(hispanic_female)
confint(hispanic_female)


### forest plot ###
#### do not use metacont() here
######## metacont() is used in raw data instead of pre-calculated effect size
##################### for pre-calculated effect size, we can use metagen()
group = "White male"
effect.size = white_male$coefficients[2]
effect.size.se <- summary(white_male)$coefficients[2,2]
white.male <- data.frame(group, effect.size, effect.size.se)
colnames(white.male) = c('group', "effect.size", "effect.size.sd")

group = "White female"
effect.size = white_female$coefficients[2]
effect.size.se <- summary(white_female)$coefficients[2,2]
white.female <- data.frame(group, effect.size, effect.size.se)
colnames(white.female) = c('group', "effect.size", "effect.size.sd")

group = "black male"
effect.size = black_male$coefficients[2]
effect.size.se <- summary(black_male)$coefficients[2,2]
black.male <- data.frame(group, effect.size, effect.size.se)
colnames(black.male) = c('group', "effect.size", "effect.size.sd")

group = "black female"
effect.size = black_female$coefficients[2]
effect.size.se <- summary(black_female)$coefficients[2,2]
black.female <- data.frame(group, effect.size, effect.size.se)
colnames(black.female) = c('group', "effect.size", "effect.size.sd")

group = "other male"
effect.size = other_male$coefficients[2]
effect.size.se <- summary(other_male)$coefficients[2,2]
other.male <- data.frame(group, effect.size, effect.size.se)
colnames(other.male) = c('group', "effect.size", "effect.size.sd")

group = "other female"
effect.size = other_female$coefficients[2]
effect.size.se <- summary(other_female)$coefficients[2,2]
other.female <- data.frame(group, effect.size, effect.size.se)
colnames(other.female) = c('group', "effect.size", "effect.size.sd")

group = "hispanic male"
effect.size = hispanic_male$coefficients[2]
effect.size.se <- summary(hispanic_male)$coefficients[2,2]
hispanic.male <- data.frame(group, effect.size, effect.size.se)
colnames(hispanic.male) = c('group', "effect.size", "effect.size.sd")

group = "hispanic female"
effect.size = hispanic_female$coefficients[2]
effect.size.se <- summary(hispanic_female)$coefficients[2,2]
hispanic.female <- data.frame(group, effect.size, effect.size.se)
colnames(hispanic.female) = c('group', "effect.size", "effect.size.sd")

meta_sex_race <- rbind(white.male, white.female, black.male, black.female,
                        hispanic.male, hispanic.female)
rownames(meta_sex_race) <- c("1", "2","3", "4", "5", "6")

meta.sex.race <- metagen(TE = effect.size,
                 seTE = effect.size.sd,
                 studlab = group,
                 data = meta_sex_race,
                 sm = "MD",
                 fixed = T,
                 random = F,
                 method.tau = "REML",
                 hakn = TRUE)

summary(meta.sex.race)

forest_sex_race = forest(meta.sex.race,print.I2 = F, print.I2.ci = F, print.tau2 = F, print.pval.Q = T, 
       leftcols = c("studlab", "effect.ci"), leftlabs = c("Strata", "Effect Estimate (95% CI)"),
       rightcols = c("w.fixed"), xlim = c(-2,2))
```

## causal mediation analysis -- physical activity

```{r}
# check the interaction between physical activity and depression #
sens_hlth <- lm(GRIM_residuals ~ depression + age + GENDER + race_ethnicity + DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO + R13ACT  , data = complete)

sens_inter_hlth <- lm(GRIM_residuals ~ depression + age + GENDER + race_ethnicity + DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO + R13ACT + R13ACT*depression  , data = complete)

summary(sens_hlth)
summary(sens_inter_hlth)

inter_act_change <- abs((summary(sens_inter_hlth)$coefficient[2,1] - summary(sens_hlth)$coefficient[2,1]))*100/summary(sens_hlth)$coefficient[2,1] ## interaction change beta 3%

# causal mediation analysis

complete2 <- complete %>% 
  mutate(high_depression = if_else(depression == "Elevated depressive symptoms", 1,0))

set.seed(2022)


med.fit <- lm(R13ACT ~ high_depression + age + GENDER + race_ethnicity + DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO, data = complete2)
summary(med.fit)
out.fit <- lm(GRIM_residuals ~ high_depression + R13ACT + age + GENDER + race_ethnicity + DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO , data = complete2)
summary(out.fit)

med.out <- mediate(med.fit, out.fit, treat = "high_depression", mediator = "R13ACT", boot = T,  sims = 1000)
summary(med.out)

## sensitivity analysis ##
sens.out <- medsens(med.out, rho.by = 0.1, effect.type = "both", sims = 100)
summary(sens.out)


### use CMAvers ##
set.seed(2022)
med_act <- cmest(data = complete, model = "rb", outcome = "GRIM_residuals", exposure = "depression",
                                mediator = "R13ACT", basec = c("age" , "GENDER" , "race_ethnicity" , "DEGREE_collapsed" , "PMARST_collapsed" , "R13CONDE" , "R13DRINKD" , "gran" , "PMONO", "smoke"), EMint = TRUE,
                                mreg = list("linear"), yreg = "linear",
                                astar = "Low or no depressive symptoms", a = "Elevated depressive symptoms", mval = list(0), 
                                estimation = "paramfunc", inference = "bootstrap", nboot = 1000)

summary(med_act)
```

### causal mediation -- smoking

```{r}
# use CMAverse #
set.seed(2022)
med_smk <- cmest(data = complete, model = "rb", outcome = "GRIM_residuals", exposure = "depression",
                                mediator = "smoke", basec = c("age" , "GENDER" , "race_ethnicity" , "DEGREE_collapsed" , "PMARST_collapsed" , "R13CONDE" , "R13DRINKD" , "gran" , "PMONO"), EMint = TRUE,
                                mreg = list("multinomial"), yreg = "linear",
                                astar = "Low or no depressive symptoms", a = "Elevated depressive symptoms", mval = list("Never"), 
                                estimation = "paramfunc", inference = "bootstrap", nboot = 1000)

summary(med_smk)
```

### causal mediation -- physical activity + smoking

```{r}
set.seed(2022)
med_smk_act <- cmest(data = complete, model = "rb", outcome = "GRIM_residuals", exposure = "depression",
                                mediator = c("smoke", "R13ACT"), basec = c("age" , "GENDER" , "race_ethnicity" , "DEGREE_collapsed" , "PMARST_collapsed" , "R13CONDE" , "R13DRINKD" , "gran" , "PMONO"), EMint = TRUE,
                                mreg = list("multinomial", "linear"), yreg = "linear",
                                astar = "Low or no depressive symptoms", a = "Elevated depressive symptoms", mval = list("Never", 0), 
                                estimation = "imputation", inference = "bootstrap", nboot = 1000)

summary(med_smk_act)
```

### multi-panel histogram of age acceleration + depression symptoms

```{r}
par(mfrow = c(3,3))
hist(complete$GRIM_residuals, xlab = "Age acceleration (Years)", main = "Histgram of GrimAge acceleration")
hist(complete$HORVATH_residuals, xlab = "Age acceleration (Years)", main = "Histgram of Horvath acceleration")
hist(complete$LEVINE_residuals, xlab = "Age acceleration (Years)", main = "Histgram of Levine acceleration")
hist(complete$MPOA_residuals, xlab = "Age acceleration (Years)", main = "Histgram of MPOA acceleration")
hist(complete$HANNUM_residuals, xlab = "Age acceleration (Years)", main = "Histgram of Hannum acceleration")
hist(complete$R13CESD, xlab = "CES-D score", main = "Histgram of CES-D score")

```

## sensitivity analysis -- excluded people older than 85 ##
```{r}
complete_85 <- complete %>% 
  filter(age <= 85)

model_crude <- lm(GRIM_residuals ~ depression, data = complete_85)

plot(model_crude)
summary(model_crude)
confint(model_crude)

model_prim <- lm(GRIM_residuals ~ depression + age + GENDER + race_ethnicity + DEGREE_collapsed + PMARST_collapsed + gran + PMONO  , data = complete_85)

plot(model_prim)
summary(model_prim)
confint(model_prim)

model_hlth <- lm(GRIM_residuals ~ depression + age + GENDER + race_ethnicity + DEGREE_collapsed + PMARST_collapsed + R13CONDE + R13DRINKD + smoke + gran + PMONO , data = complete_85)

plot(model_hlth)
summary(model_hlth)
confint(model_hlth)
```

```{r}

```

